\subsection{Business 2}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Red{die Firma} & firm/enterprise & \Red{die Bezahlung} & salary \\
	\Red{die Lieferung} & delivery & das Lager & storage/warehouse \\
	\Red{die Industrie} & industry & \Red{die Gr{\"u}ndung} & foundation \\
	\Red{die Ware} & merchandise & \Red{die Ausgabe} & edition/version \\
	\Red{die Tabelle} & chart/table & \Red{die B{\"o}rse} & stock market \\
	\Red{die Versicherung} & insurance & \Red{die Produktion} & production \\
	\Red{die Garantie} & guarantee & \Red{die Leistung} & achievement \\
	\Red{die Anzeige} & advertisement & \Blue{der Kundenservice} & customer service \\
	\Red{die Werbung} & advertising [effort] & \Red{die Branche} & branch \\
	das Angebot(e) & deal(s) & \Red{die Logistik} & logistics \\
	\Red{die Chance} & chance & \Blue{der Bedarf} & requirement \\
	\Blue{der Antrag} & application & \Blue{der Betrieb} & business \\
	\Blue{der Auftrag} & task & verhandelen & to negotiate \\
	\Blue{der Service} & service & \Blue{der Einzelhandel} & retail sector \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Sein Haus steht zum \Blue{Verkauf}. \\
  His house is for \Blue{sale}.
  \item  Er ist hoch oben in der Firma. \\
  He is at the top at the company.
  \item  Die \Red{Messe} war sch{\"o}n. \\
  The \Red{trade fair} was nice.
  \item  Was ist die \Red{Marke}? \\
  What is the \Red{brand}?
  \item  Ich teile deine \Red{Bewertung}. \\
  I share your \Red{assessment}.
  \item  Wir haben ein \Blue{Stellenangebot} in der Zeitung. \\
  We have a \Blue{job offer} in the newspaper (``position'' + ``deal'').
  \item  Was f{\"u}r \Red{Deinstleistungen} haben Sie? \\
  What do you have for \Red{services}?
\end{itemize}


\pagebreak
\subsection{Future 4}

\begin{itemize}
  \item  Wir werden ein neues Fenster \textbf{ausw{\"a}hlen}. \\
  We will \textbf{choose} a new window.
  \item  Wann werdet ihr das Datum \textbf{festlegen}? \\
  When are you going to \textbf{determine} the date?
  \item  Ich werde f{\"u}r Sie {\"u}bernehmen. \\
  I wll take over for you.
  \item  Wir werden mehr \textbf{leisten}. \\
  We will \textbf{accomplish} more.
  \item  Ich werde mitmachen. \\
  I will partake.
  \item  Du wirst mich \textbf{ansehen}. \\
  You will look at me.
  \item  Ich werde alles erf{\"u}llen. \\
  I will accomplish everything.
  \item  Wir werden sparen m{\"u}ssen. \\
  We will have to economize.
  \item  Ich werde den Computer \textbf{entfernen}. \\
  I will \textbf{remove} the computer.
  \item  Ich werde f{\"u}r dich sorgen! \\
  I will take care of you!
  \item  Ich werde Sie \textbf{anzeigen}! \\
  I will \textbf{report} you!
\end{itemize}


\pagebreak
\subsection{Sports}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	Fu{\ss}ball & soccer & \Red{die Mannschaft} & team (professional?) \\
	\Blue{der Ball} & ball & \Blue{der Spieler}, \Red{die Spielerin} & player \\
	\Blue{der Sport} & sport & das Spiel & match/game/play \\
	das Hobby & hobby & \Blue{der Zuschauer}, \Red{die Zuschauerin} & spectator \\
	\Red{die Aktivit{\"a}t} & activity & \Red{die Liga} & league \\
	das Tennis & tennis & \Blue{der Ski} & ski \\
	das Team & team & das Schwimmbad & swimming pool \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Der \Blue{Sieg} ist unser! \\
  \Blue{Victory} is ours!
  \item  Eine \Red{Teilnahme} ist kostenlos. \\
  \Red{Participation} is free.
  \item  Wir m{\"o}gen die \textbf{Bundesliga}. \\
  We like the \textbf{German soccer league}.
  \item  Die \Red{Wette} gilt! \\
  The \Red{bet} is valid!
  \item  Was machst du gerne in deiner \Red{Freizeit}? \\
  What do you like to do in your \Red{free time}?
  \item  Der Vogel ist meist in \Red{Bewegung}. \\
  The bird is often in \Red{motion}.
  \item  Freizeit ist Kindern wichtig. \\
  Free time is important to children.
\end{itemize}


\pagebreak
\subsection{Perfect 2}

\begin{itemize}
  \item  Ich habe dich nicht \textbf{verstanden}. \\
  I did not understand you.
  \item  Ist etwas passiert? \\
  Did something happen?
  \item  Sie hat an dich \textbf{gedacht}. \\
  She was \textbf{thinking} about you.
  \item  Das habe ich auch gedacht. \\
  I also have thought of this.
  \item  Ich wei{\ss} nicht genau, wie es passiert ist. \\
  I do not exactly know how it happened.
  \item  Haben Sie verstanden? \\
  Do you understand?
  \item  Ich habe verstanden? \\
  I understood.
  \item  Ist gestern was passiert? \\
  What happened yesterday?
  \item  Hast du es \textbf{gebracht}? \\
  Have you brought it?
  \item  Sie hat eine Firma \textbf{gegr{\"u}ndet}. \\
  She has \textbf{founded} a company.
  \item  Er hat das \textbf{gewusst}. \\
  He \textbf{knew} it.
  \item  Sie hat das B{\"u}ro schon verlassen. \\
  She has already left the office.
  \item  Keiner hat mich je so genannt. \\
  Nobody has ever called me that.
  \item  Meine Pflanze ist \textbf{verschwunden}. \\
  My plant is \textbf{gone}.
  \item  Silber und Gold have ich nie \textbf{gewollt}. \\
  I have never \textbf{wanted} silver and gold.
  \item  Sie haben \textbf{gew{\"a}hlt}. \\
  They have \textbf{voted}.
  \item  Ich habe deinen Wunsch \textbf{erf{\"u}llt}. \\
  I have \textbf{fulfilled} your wish.
\end{itemize}


\pagebreak
\subsection{Art}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Red{die Kunst, die K{\"u}nste} & art(s) & tanzen & to dance \\
	\Blue{K{\"u}nstler}, \Red{K{\"u}nstlerin} & artists & \Red{die Literatur} & literature \\
	das Studio & studio & das Museum & musuem \\
	\Red{die Galerie} & gallery & \Red{die Sammlung} & collection \\
	\Red{die Ausstellung} & exhibition & \Red{die Fotografie} & photography \\
	\Red{die Gitarre} & guitar & das Design & design \\
	\Red{die Musik} & music & \Red{die Kritik} & critique/criticism \\
	\Blue{der Musiker}, \Red{die Musikerin} & musician & \Red{die Mode} & fashion \\
	das Theater & theater & \Red{die B{\"u}hne} & stage \\
	\Red{die Kultur} & culture & \Blue{der Rahmen} & frame/framework \\
	malen & to paint & \Red{die Kamera} & camera \\
	das Instrument(e) & instrument(s) & schauen & to look \\
	das Musikinstrument(e) & musical instrument(s) & \Blue{der Schauspieler}, \Red{die Schauspielerin} & actor, actress \\
	das Modell & model & das Kino & cinema/movie theater \\
	\Blue{der Tanz} & dance \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Mein \Blue{Entwurf} war beliebt. \\
  My \Blue{design} was popular.
  \item  Der \Blue{Film} ist langweilig. \\
  The \Blue{movie} is boring.
  \item  Alles ist f{\"u}r den \Blue{Start} bereit. \\
  Everything is ready for the \Blue{start}.
  \item  Er mag \Blue{Jazz} und ich auch. \\
  He likes \Blue{Jazz} and I do too.
  \item  Die \Red{Konzerte} sind im Schloss. \\
  The \Red{concerts} are in the castle.
  \item  Er hat seinen eigenen \Blue{Stil}. \\
  He has his own \Blue{style}.
  \item  Die \Red{Dokumentation} stimmt. \\
  The \Red{documentary} is accurate.
%  \item  Er begann verschiedene \Red{T{\"a}tigkeiten}. \\
%  He began various activities.
  \item  Regie! \\
  Direction!\footnote{"From my experience in the theater, I can think of two instances to use this depending on the intended meaning.  The first is during a standing ovation (or other great appreciation) at the end of a play. The cast will often come out for a bow, and then bring out the director. Sometimes the audience will also call, "Author!" (The author of the play.) I can see the audience calling for the director instead if he does not come out on his own.  The second time is if the call is for "direction," as opposed to "director." Sometimes an actor will say, "Line?" if he needs prompting. In the same vein, if a rehearsal is not going well or the director isn't actively participating, the actors might call for direction." -- Duolingo user Immolatrix.}
\end{itemize}


\pagebreak
\subsection{Passive Voice}

\begin{itemize}
  \item  Der Lehrer wurde bewertet. \\
  The teacher was evaluated.
  \item  Dieses Wort wird geschrieben. \\
  This word gets written.
  \item  Und hier wurde er \textbf{geboren}. \\
  And here he was \textbf{born}.
  \item  Die Bewerbung wurde nicht \textbf{ber{\"u}chsichtigt}. \\
  The application was not considered.
  \item  Beide wurden bereits zuvor ver{\"o}ffenlicht. \\
  Both were already published before.
  \item  Hilfe, ich werde gegessen! \\
  Help, I am being eaten!
  \item  Das alte Buch wird auch heute noch genutzt. \\
  The old book is still used today.
  \item  Mitglieder werden bevorzugt. \\
  Members are preferred.
  \item  Wird das gew{\"a}hrleistet? 
  Is that guaranteed?
  \item  Es wurde ge{\"a}ndert. \\
  It was changed.
  \item  Es wird wie eine Suppe gegessen. \\
  It gets eaten like a soup.
  \item  Fleisch wird selten gegessen. \\
  Meat is rarely eaten.
\end{itemize}


\pagebreak
\subsection{Religion}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Blue{der Gott} & god & \Blue{Atheist}, \Red{Atheistin} & atheist \\
	\Blue{der Glaube} & belief & \Blue{der Hinduismus} & Hinduism \\
	\Blue{der Priester}, \Red{die Priesterin} & priest, priestess & \Blue{der Buddhismus} & Buddhism \\
	beten & to pray & heilig & holy \\
	das Gebet & prayer & \Red{die Kirche} & church \\
	das Paradies & paradise & \Blue{der M{\"o}nch(e)}, \Red{die M{\"o}nchin} & monk(s) \\
	\Blue{Christ}, \Red{Christin} & Christian & meditieren & to meditate \\
	\Blue{Muslim}, \Red{Muslimin} & Muslim & \Blue{der Tempel} & temple \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Wir reden {\"u}ber \Red{Religion}. \\
  We speak about \Red{religion}.
  \item  Der \Blue{Tod} hat das letzte Wort. \\
  \Blue{Death} has the last word.
  \item  Die Welt \Blue{Gottes} ist gro{\ss}. \\
  The world of \Blue{God} is large.
  \item  Wohin werden wir nach dem Tod gehen? \\
  Where do we go after death?
  \item  Sie glauben an die Worte Gottes. \\
  You believe in the words of God.
  \item  Was wei{\ss}t du {\"u}ber den \Blue{Islam}? \\
  What do you know about \Blue{Islam}?
  \item  Glaubst du an \Red{Wiedergeburt}? \\
  Do you believe in \Red{reincarnation}?
  \item  Christen, Muslime, und Juden leben hier. \\
  Christians, Muslims, and Jews live here.
  \item  Mit diesem Gebet lernt er Franz{\"o}sisch. \\
  With this prayer, he learns French.
  \item  Die \Red{Synagoge} ist neben der \Red{Moschee}. \\
  The synagogue is adjacent to the mosque.
\end{itemize}


\pagebreak
\subsection{Politics}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Red{die Politik} & politics & das Ministerium & ministry \\
	\Red{die Stimme} & vote/voice & das Recht & right \\
	\Red{die Partei} & party & \Red{die Genehmigung} & permission/approval \\
	das Interesse & interest & \Blue{der Prinz}, \Red{die Prinzessin} & prince, princess \\
	\Blue{der Politiker}, \Red{die Politikerin} & politician & \Blue{der Einfluss} & influence \\
	\Red{die Republik} & republic & \Red{die Regel} & rule \\
	\Red{die Vertrag, die Vertr{\"a}ge} & contract(s) & \Red{die Vorschrift} & regulation/rule/instruction \\
	\Blue{der Erfolg} & success & das Rathaus & city hall \\
	\Blue{der K{\"o}nig}, \Red{die K{\"o}nigin} & king, queen & w{\"a}hlen & to vote \\
	\Blue{der Staat} & state & \Blue{der Schaden, die Sch{\"a}den} & damage(s) \\
	\Red{die Verwaltung} & administration & das Verfahren & trial/process \\
	das Gesetz(e) & law(s) & \Blue{der Prozess} & trial/process \\
	\Red{die Beh{\"o}rde} & authority/agency & \Red{die Entscheidung} & verdict/decision \\
	\Red{die Regierung} & government & \Red{die Bedingung} & requirement \\
	\Blue{der Kampf} & struggle \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Der \Blue{Pr{\"a}sident} hat mehr Macht als der B{\"u}rgermeister. \\
  The \Blue{president} has more power than the mayor.
  \item  Wir sind das \textbf{Volk}. \\
  We are the people.\footnote{\url{https://www.thoughtco.com/people-leute-menschen-volk-4069439}}
  \item  Heute sind \textbf{Wahlen}. \\
  \textbf{Elections} are today.
  \item  \Blue{Krieg} und \Blue{Frieden} \\
  War and peace
  \item  Was h{\"a}ltst du von dem Krieg? \\
  What do you think of the war?
  \item  Die Lage ist unter \Red{Kontrolle}. \\
  The situation is under control.
\end{itemize}


\pagebreak
\subsection{Adverbs 4}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	jederzeit & any time & hingegen & on the other hand \\
	schlie{\ss}lich & finally & weiterhin & still/continue to \\
	dennach & nevertheless & derzeit & currently \\
	mittlerweile & by now & insbesondere & particularly \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Er ist \textbf{ebenfalls} aktiv. \\
  He is active \textbf{as well}.
  \item  Ich sah es \textbf{soeben}. \\
  I \textbf{just} saw it.
  \item  Wer \textbf{kommt} noch \textbf{hinzu}? \\
  Who \textbf{else is coming}?
  \item  Bist du \textbf{hierzu} bereit? \\
  Are you ready \textbf{for this}?
  \item  \textbf{Deswegen} ist der Saft super. \\
  \textbf{That is why} the juice is great.
  \item  \textbf{Hiermit} m{\"o}chte ich bezahlen. \\
  I would like to pay \textbf{with this}.
  \item  Der Autor liest das Buch \textbf{nochmals}. \\
  The author is reading the book \textbf{again}.
\end{itemize}


\pagebreak
\subsection{Abstract Objects 3}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Blue{der Vergleich} & comparison & \Red{die Vermittlung} & agency/negotiation \\
	das System(e) & system(s) & \Blue{der Umfang} & perimeter/circumference \\
	\Red{die Anwendung} & application (of an idea) & \Red{die Verpflichtung} & responsibility/obligation \\
	\Red{die F{\"o}rderung} & funding & \Red{die Oberfl{\"a}che} & surface [also: surface area] \\
	das Kriterion, die Kriterien & criterion, criteria & \Red{die Vorstellung} & notion/idea \\
	\Red{die Anlag} & asset & \Blue{der Vortrag} & lecture \\
	\Red{die Rubrik} & rubric & \Red{die Gewalt} & violence \\
	\Blue{der Durchschnitt} & average (``by'' + ``cut'') & \Red{die Gefahr} & danger/threat \\
	\Blue{der Abschnitt} & segment (``from'' + ``cut'') & \Blue{der Charakter(e)} & characters \\
	\Blue{der Verlust} & loss & das Material(ien) & material(s) \\
	das Beispiel & example & \Red{die Qualifikation} & qualification \\
	\Blue{der Eintritt} & admission & \Blue{der Bezug} & salary/cover \\
	\Red{die Er{\"o}ffnung} & opening & \Red{die Runde} & lap/circle/turn \\
	\Blue{der Inhalt(e)} & content(s) & \Blue{der Eingang} & entrance \\
	\Red{die Reparatur} & repair & \Blue{der Abschluss} & graduation/degree (``from'' + ``ending'') \\
	\Red{die Geschwindigkeit} & speed & \Red{die Einstellung} & setting/attitude/adjustment \\
	\Blue{der Rest} & rest/remainder & \Blue{der Empfang} & reception \\
	\Blue{der Anteil} & share & \Blue{der Ansatz} & approach \\
	\Red{die Versorgung} & supply & \Red{die Karte} & map/ticket \\
	\Red{die Auflage} & output/circulation/edition & \Red{die Kooperation} & cooperation \\
	das Ereignis(se) & occasion(s) & \Red{die Schau} & show/exhibition \\
	\Red{die Tradition} & tradition & das Symbol & symbol \\
	das Niveau & level & \Blue{der F{\"a}ll(e)} & case(s)/circumstance(s) \\
	\Red{die Agentur} & agency \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Mit welcher Geschwindigkeit l{\"a}uft er? \\
  How fast is he running?
  \item  Im \textbf{Umgang} mit Kindern \\
  in \textbf{dealing} with children
\end{itemize}


\pagebreak
\subsection{Verbs: Conditional 2}

\begin{itemize}
  \item  Ich w{\"u}rde gern mit dir springen. \\
  I would like to jump with you.
  \item  Wann w{\"u}rde man hier wandern? \\
  When would one go hiking here?
  \item  Er w{\"u}rde die Post f{\"u}r mich sammeln? \\
  He would collect the mail for me.
  \item  Ich w{\"u}rde ein Bier lieben. \\
  I would love a beer.
  \item  Er w{\"u}rde das nicht vorschlagen. \\
  He would not suggest that.
  \item  Sie w{\"u}rde viel mehr leisten. \\
  She would achieve much more.
  \item  W{\"u}rdest du mir meine Frage beantworten? \\
  Would you answer my question?
  \item  Wer w{\"u}rde darum bitten? \\
  Who would ask for that?
  \item  Wer w{\"u}rde mich \textbf{darstellen}? \\
  Who would \textbf{play} me?
  \item  Das w{\"u}rde mich freuen. \\
  That would make me happy.
  \item  Ich w{\"u}rde \textbf{vorsichtig} sein. \\
  I would be \textbf{careful}.
\end{itemize}


\pagebreak
\subsection{Philosophy}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Blue{der Pessimist} & pessimist & Philosophie & philosophy \\
	\Blue{der Optimist} & optimist & \Red{die Wirklichkeit} & reality \\
	skeptisch & skeptical & das Bewusstsein & consciousness \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Die Philosophie passt zu dir. \\
  The philosophy suits you.
  \item  Sie ist \textbf{bei Bewusstsein}. \\
  She is [medically] \textbf{conscious}.
  \item  Wahr oder falsch? \\
  True or false?
\end{itemize}


\pagebreak
\subsection{Present 4}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	f{\"u}hren & to lead & erf{\"a}hren & to hear/learn \\
	empfehlen & to recommend & garantieren & to guarantee \\
	versuchen & to try & l{\"a}ssen & to leave (behind) \\
	existieren & to exist & aktivieren & to activate \\
	schaffen & to create & vertriten & to substitute (for) \\
	fordern & to demand & h{\"a}ngen & to hang \\
	entstehen & to arise & bilden & to form \\
	l{\"o}sen & to solve/loosen & vergiben & to forgive \\
	erkl{\"a}ren & to explain & liefern & to deliver \\
	er{\"o}ffnen & to open up & leisten & to execute \\
	ben{\"o}tigen & to need (urgently) & behandelen & to cover/treat \\
	entfernen & to remove & berechtigen & to authorize \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Wie \textbf{wirkt} sie auf dich? \\
  How does she \textbf{affect} you?
%  \item  Bitte diesen Hinweis beachten. \\
%  Please consider this tip.
  \item  Wir \textbf{verf{\"u}gen {\"u}ber} viel Geld. \\
  We \textbf{have} a lot of money.
  \item  Wir verlassen das Land f{\"u}r immer. \\
  We are leaving the country forever.
  \item  Ich wei{\ss} nicht, warum wir dar{\"u}ber diskutieren. \\
  I do not know why we are discussing this.
  \item  Was soll das? \\
  What is the point?
  \item  Ich kann mir das nicht leisten. \\
  I cannot afford that.
  \item  Zwei Eier \textbf{gen{\"u}gen}. \\
  Two eggs are enough.
  \item  Keine Angst, wir planen nichts. \\
  Do not worry, we are planning nothing.
  \item  Dann planen wir es gemeinsam. \\
  In that case, we plan it together.
  \item  Das h{\"a}ngt von ihnen ab. \\
  That depends on you.
  \item  Es \textbf{besteht} aus Holz. \\
  It is \textbf{made} out of wood.
  \item  Wir \textbf{bearbeiten} das Buch. \\
  We \textbf{are working on} the book.
\end{itemize}


\pagebreak
\subsection{Fantasy and Science Fiction}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Blue{der Planet} & planet & \Red{die Magie} & magic \\
	retten & to rescue & das Einhorn & unicorn \\
%	entdecken & to discover \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Die \Red{Hexe} hat den Prinzen in einen \Blue{Drachen} \textbf{verwandelt}. \\
  The \Red{witch} has \textbf{transformed} the prince into a \Blue{dragon}.
  \item  Eine junge \Red{Galaxie} wurde \textbf{entdeckt}. \\
  A young \Red{galaxy} was \textbf{discovered}.
  \item  Die \Red{Au{\ss}erirdische} hat ein wei{\ss}es \textbf{Raumschiff}. \\
  The \Red{alien} has a white \textbf{spaceship}.
  \item  ``Der \Blue{Weltraum}:  Unendliche Weiten'' \\
  ``Space: the final frontier''
  \item  Sie kommt aus einer anderen \Red{Dimension}. \\
  She comes from another \Red{dimension}.
  \item  \textbf{Verschwinden} wir von hier! \\
  Let's \textbf{get out} of here!
  \item  Sie verschwindet! \\
  She is disappearing!
  \item  Das ist \textbf{unglaublich}! \\
  That is \textbf{incredible}!
  \item  Ich will eine \Red{Zeitreise} machen! \\
  I will travel through \Red{time}!
  \item  Er w{\"u}rde mit der \Blue{Zeitmaschine} in die Vergangenheit reisen. \\
  He would travel to the past with the \Blue{time machine}.
  \item  Es war eine \textbf{magische} Nacht. \\
  It was a \textbf{magical} night.
\end{itemize}


\pagebreak
\subsection{Abstract Objects 4}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Red{die Berechnung} & calculation & \Red{die Aufnahme} & record/recording/admission \\
	das Prinzip & principle & \Red{die Angab} & specification/information \\
	\Blue{der Sitz} & seat/headquarters & \Red{die Aufgab} & task/duty \\
	\Red{die Option} & option & \Red{die Hausaufgaben} & homework \\
	\Red{die Figur} & figure & \Blue{der Punkt(e)} & point(s) \\
	\Red{die Herkunft} & origin & \Red{die Ebene} & level \\
	\Blue{der Anschluss} & connection & \Blue{der Ausdruck} & expression \\
	\Red{die Serie} & series & das Risiko & risk \\
	\Blue{der Ton} & note/sound/clay & \Red{die Bezeichnung} & label/designation/name \\
	\Red{die Abteilung} & department & \Red{die Struktur} & structure \\
	\Red{die Einheit} & unit & \Blue{der Schutz} & cover/protection/guard \\
	\Red{die Austauch} & exchange & \Blue{der Schritt} & step/pace \\
	das Vorbereitung & preparation & \Red{die Folge} & sequence \\
	\Red{die Wertung} & rating & \Red{die Verantwortung} & responsibility \\
	\Blue{der Gegensatz} & conflict/contrast (``against'' + ``sentence'') & \Blue{der Standard} & standard(s) \\
	\Red{die Fassung} & version/composure & \Red{die M{\"o}glichkeit} & possibility \\
	\Blue{der Vorteil} & advantage & \Red{die Tiefe} & depth/bass \\
	\Red{die Betreuung} & supervision/care & \Red{die Breite} & width/spread/latitude \\
	\Red{die Sicht} & point of view & \Blue{der Wechsel} & change/transition \\
	\Red{die Ansicht} & opinion & \Red{die Anforderung} & demand \\
	\Red{die {\"U}bersicht} & overview & das Detail(s) & detail(s) \\
	\Red{die Quell} & source & \Red{die Art} & species/kind \\
	das Format & format & \Red{die Grundlage} & basis/foundation \\
	\Red{die Situation} & situation & \Red{die Alternative} & alternative \\
	\Blue{der Zusammenhang} & context (``together'' + ``slope'') & \Red{die Mitte} & middle \\
	\Red{die Leitung} & leadership & das Teil & section/part \\
	\Red{die Linie} & line & \Red{die Funktion} & function \\
	\Red{die Entfernung} & distance & \Blue{der Fehler} & mistake \\
	\Red{die Ausnahme} & exception & \Red{die Anregungen} & suggestion \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Wir denken nie an diese \Red{Merkmale}. \\
  We never think about these \Red{features}.
  \item  Leben ist der \Blue{Zweck} und ist sinnvoll. \\
  Life is the \Blue{purpose} and is meaningful.
  \item  Das ist der \Blue{Lauf} der Geschichte. \\
  That is the \Blue{course} of history. 
  \item  Der Weg zur deutschen Einheit \\
  the path to German reunification.
  \item  Wer ist an der Reihe? \\
  Whose turn is it? [formal version of ``Wer ist dran?'']
  \item  Das \textbf{Band} ist l{\"a}nger als das Bein. \\
  The \textbf{band} is longer than the leg.
\end{itemize}

Note that \Red{die Nutzung} is the usage of something while \Red{die Nutzen} is the benefit.


\pagebreak
\subsection{Relative Pronouns}

The relative pronouns follow the same pattern as the definite articles with the exception of the dative plural and the genitive forms.  In German, relative clauses are always set off by commas from the rest of the sentence.  There's no distinction between restrictive and non-restrictive clauses.  Also, relative pronouns can never be dropped.  The form you need to use is governed by the grammatical gender and number of the word that is being referred to (outside the relative clause), and the case is governed by the context of the relative clause.  Keep in mind that certain prepositions and verbs always trigger a certain case, e.g. the preposition "mit" always takes the dative case and so does the verb "helfen".

\begin{center}\begin{tabular}{c|c|c|c|c}
  ~ & masculine & feminine & neuter & plural \\
  \hline
  nominative & der & die & das & die \\
  accusative & den & die & das & die \\
  dative & dem & der & dem & \textbf{denen} \\
  genitive & \textbf{dessen} & \textbf{deren} & \textbf{dessen} & \textbf{deren} \\
\end{tabular}\end{center}

\begin{itemize}
%  \item  Das ist ein Name, der sehr alt ist. \\
%  That is a name which is very old.
%  \item  Es ist ein Tier, das einen langen Hals hat. \\
%  It is an animal that has a long neck.
  \item  Wir sind Studenten, die gerne Bier trinken. \\
  We are students that like to drink beer.
  \item  Sie ist eine Frau, die wei{\ss}, was sie will. \\
  She is a woman that knows what she wants.
  \item  Habt ihr etwas, das \textbf{billiger} ist? \\
  Do you have something that is \textbf{cheaper}?
  \item  Ist das die gr{\"o}{\ss}te Wohnung, die Sie haben? \\
  Is that the largest apartment that you have?
  \item  Sie ist die Frau, die ich liebe. \\
  She is the woman that I love.
  \item  Ist sie die Frau, die du liebst? \\
  Is she the woman that you love?
  \item  Er ist ein Politiker, \textit{dem} ich glaube. \\
  He is a politician that I believe.
  \item  Das Dorf, in dem ich wohne, ist klein. \\
  The village in which I live is small.
  \item  Die H{\"a}user, in \textbf{denen} wir wohnen, sind alt. \\
  The houses in which we live are old.
  \item  Das ist die Frau, der ich \textbf{geholfen} habe. \\
  That is the woman that I have \textbf{helped}.
  \item  Es ist ein Pferd, \textbf{dessen} Namen ich nicht kenne. \\
  It is a horse \textbf{whose} name I do not know.
  \item  Das ist meine Schwester, deren Freund Lehrer ist. \\
  This is my sister whose boyfriend is a teacher.
  \item  Das sind Kinder, deren Eltern reich sind. \\
  Those are children whose parents are rich.
\end{itemize}


\pagebreak
\subsection{Classical Music}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Red{die Geige} & violin & dirigieren & to conduct/direct \\
	\Blue{der Komponist}, \Red{die Komponistin} & composer & \Blue{der Dirigent}, \Red{die Dirigentin} & conductor/director \\
	das Klavier & piano & \Red{die Oper} & opera \\
	das Orchester & orchestra & \Blue{der S{\"a}nger}, \Red{die S{\"a}ngerin} & singer \\
	\Red{die Trommel} & drum & klingen & to sound \\
\end{tabular}\end{center}

\begin{itemize}
  \item  Sie spielt eine sch{\"o}ne \Red{Melodie} auf der \Red{Fl{\"o}te}. \\
  She is playing a beautiful \Red{melody} on the \Red{flute}.
  \item  Dieses Orchestra hat einen bekannten Dirigenten. \\
  This orchestra has a well-known conductor.
\end{itemize}


\pagebreak
\subsection{Politics 2}

\begin{center}\begin{tabular}{r|l||r|l}
  \textbf{Deutsch} & \textbf{English} & \textbf{Deutsch} & \textbf{English} \\
	\hline
	\Red{die Freiheit} & freedom & \Red{die Vereinbarung} & agreement \\
	\Red{die Strategie} & strategy & \Blue{der Rat} & advice/council \\
	\Red{die Veranstaltung} & event & korrupt & corrupt \\
	\Red{die Bundesregierung} & federal government & \Blue{der Verband} & organization/alliance \\
	\Red{die Polizei} & police & \Blue{der Anspr{\"u}ch(e)} & expectation(s) \\
	\Red{die Steuer} & tax & protestieren & to protest \\
	\Red{die Umfrage} & survey/poll & \Red{die Wirtschaft} & economy \\
\end{tabular}\end{center}

\begin{itemize}
  \item  die \textbf{Bundesrepublik} Deutschland \\
  the German \textbf{Federal Republic}
  \item  Stimmst du dem \textbf{Urteil} zu? \\
  Do you agree with the \textbf{verdict}?
  \item  Das Reich steht auf dem Spiel. \\
  The empire is at stake.
  \item  Welches \textbf{Amt} ist zust{\"a}ndig? \\
  Which [governmental] \textbf{department} is in charge?
\end{itemize}


\pagebreak
\subsection{German Culture}

A \Red{Wurst} is a sausage. It does not specifically refer to any kind of sausage. It could be a salami, chorizo, mortadella, frankfurter, etc.  \Red{Bratwurst} specifically refers to a fried or grilled sausage.

\begin{itemize}
  \item  Wir essen \textbf{Sauerkraut} und trinken \textbf{Pils}. \\
  We are eating sauerkraut and drinking pilsner.
  \item  Is das Fest im Juli? \\
  Is the festival in July?
  \item  Wie viele \Red{W{\"u}rste} habt ihr gegessen? \\
  How many \Red{sausages} have you eaten?
  \item  Das \textbf{Oktoberfest} beginnt im September. \\
  \textbf{Oktoberfest} begins in September.
  \item  Jegen Mogen esse ich eine \Red{Brezel}. \\
  Every morning I eat a pretzel.
  \item  In Frankreich gibt es kein Oktoberfest. \\
  In France there is no Oktoberfest.
  \item  Wir \textbf{ferien} den Tag der Deutschen Einheit. \\
  We are \textbf{celebrating} German Unity Day.
  \item  Arbeitet ihr in Berlin? \\
  Do you work in Berlin?
  \item  Er ist zum \Blue{Kiosk} gegangen. \\
  He went to the \Blue{corner store}.
  \item  Feiert ihr Weihnachten? \\
  Do you celebrate Christmas?
\end{itemize}


\pagebreak
\subsection{The World}

\begin{itemize}
  \item  Ist Australien ein \Blue{Kontinent}? \\
  Is Australia a \Blue{continent}?
  \item  Ich lebe in der \Blue{Europ{\"a}ischen Union}? \\
  I live in the \Blue{European Union}.
  \item  Er ist \textbf{europ{\"a}isch}. \\
  He is \textbf{European}.
  \item  Bist du in China? \\
  Are you in China?
  \item  Indien hat viele alte Traditionen. \\
  India has many ancient traditions.
  \item  Warst du schon mal in China? \\
  Have you ever been to China?
  \item  In Asien gibt es auch \Red{Pyramiden}. \\
  There are \Red{pyramids} in Asia too.
  \item  Ich komme aus S{\"u}damerika, aber lebe in Nordamerika. \\
  I am coming from South America, but I live in North America.
  \item  {\"A}gypten ist ein Land in Afrika. \\
  Egypt is a country in Africa.
  \item  Die Wissenschaftlerin hat eine \Red{Pyramide} in S{\"u}damerika entdeckt. \\
  The scientist has discovered a \Red{pyramid} in South America.
  \item  Die Erde hat einen \Blue{Nordpol} und einen \Blue{S{\"u}dpol}.
  The Earth has a North Pole and a South Pole.
  \item  Ich bin aus der T{\"u}rkei. \\
  I am from Turkey.
  \item  Sie kommt aus Russland. \\
  She comes from Russia.
  \item  Polen und Deutschland sind Nachbarn. \\
  Poland and Germany are neighbors.
  \item  Er m{\"o}chte zum S{\"u}dpol reisen. \\
  He wants to travel to the South Pole.
\end{itemize}